---
title: "Lineage distinction analysis"
author: "Maisy Hallam"
date: "2023-04-14"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) # for data preprocessing
library(ggplot2) # for plotting
library(lme4) # for model fitting
```

# Preprocessing the data

## Import the data

First, we import all the raw data.

```{r import}
lineage_prop_files <- list.files(path="../../data/raw/lineage/prop/",pattern="*.csv",full.names=TRUE)
lineage_data <- do.call(rbind, lapply(lineage_prop_files, function(x) read.csv(x)))
```


## Process the data

We add a column TYPE which tells us if the data comes from kinbank or if it is simulated, and we delete the first two columns because I am an idiot who ran my include_headings function too many times.

```{r}
lineage_data <- lineage_data %>%
  mutate(
    
    TYPE = ifelse (grepl('frankenlanguage', LANGUAGE),
    'simulation',
    'kinbank'
  ))


lineage_data$GEN_0 <- as.numeric(lineage_data$GEN_0)
lineage_data$GEN_1 <- as.numeric(lineage_data$GEN_1)


lineage_data
```

And then we save lineage_data to a new CSV file.

```{r}
write.csv(lineage_data, '../../data/processed/all_lineage_proportions.csv')
```

Now, let's split `lineage_data` into two dataframes - `kinbank` and `simulations`.

```{r}
simulations <- filter(lineage_data,TYPE == 'simulation')
kinbank <- filter(lineage_data,TYPE == 'kinbank')
```

And save those too.

```{r}
write.csv(kinbank, '../../data/processed/kinbank_lineage_proportions.csv')
write.csv(simulations, '../../data/processed/simulated_lineage_proportions.csv')

```

# Analysis

We need the means of each SET of data - so all 1000 simulations, plus the kinbank data.

```{r}
simulation_means <- group_by(simulations,SET) %>%
  summarise(
    MEAN0 = mean(GEN_0),
    MEAN1 = mean(GEN_1)
  )

kinbank_mean <- group_by(kinbank,SET) %>%
  summarise(
    MEAN0 = mean(GEN_0),
    MEAN1 = mean(GEN_1)
  )

simulation_means
kinbank_mean
```

To plot the data, let's first take a sample of the data.

```{r}
sim_prop_sample <- sample_n(simulations,767, replace=FALSE)
kinbank_prop_sample <- sample_n(kinbank,767,replace=FALSE)
```

And we'll plot some histograms, so we can sanity check that we've done the simulations right.

```{r}
rbind(sim_prop_sample,kinbank_prop_sample) %>%
  ggplot() +
  geom_histogram(aes(GEN_0,fill=TYPE),position='dodge')

```

```{r}
simulation_means %>%
  ggplot(aes(MEAN0)) +
  geom_histogram(fill = 'lightblue', color = 'navy', alpha = 1) +
  geom_vline(aes(xintercept = 0.5077), color="navy", linetype="dashed", size=0.5) +
  labs(
    title = 'Mean proportion of lineage distinctions in Ego generation',
    color = legend) +
  xlab('GEN_0 means')
```


```{r}
rbind(sim_prop_sample,kinbank_prop_sample) %>%
  ggplot() +
  geom_histogram(aes(GEN_1,fill=TYPE),position='dodge')

```

```{r}
simulation_means %>%
  ggplot(aes(MEAN1)) +
  geom_histogram(fill = 'lightblue', color = 'navy', alpha = 1) +
  geom_vline(aes(xintercept = 0.6852), color="navy", linetype="dashed", size=0.5) +
  labs(
    title = 'Mean proportion of lineage distinctions in Ego parent generation',
    color = legend) +
  xlab('GEN_1 means')
```

Looks like the simulations converge to the real data! Phew!

## Real serious plots

First, let's plot the languages with GEN_0 proportion of lineage distinctions on the x axis, and GEN_1 proportion on the y axis.

```{r}
rbind(sim_prop_sample,kinbank_prop_sample) %>%
  ggplot(aes(x=GEN_0, y=GEN_1,color=TYPE)) +
  geom_jitter(alpha=0.2,width = 0.2) +
  xlim(0,1) +
  ylim(0,1) +
  geom_smooth(method=lm)
```

## Super serious analysis

```{r}
cor.test(simulations$GEN_0,simulations$GEN_1, method = 'spearman')
```

```{r}
cor.test(kinbank$GEN_0,kinbank$GEN_1, method = 'spearman')

```
