---
title: "What's wrong with my sims :("
author: "Maisy Hallam"
date: "2023-03-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) # for data preprocessing
library(ggplot2) # for plotting
library(lme4) # for model fitting
```

# Preprocessing the data

## Import the data

First, we import all the raw data.

```{r import}
gender_prop_files <- list.files(path="../../data/raw/gender/prop/",pattern="*.csv",full.names=TRUE)
gender_data <- do.call(rbind, lapply(gender_prop_files, function(x) read.csv(x)))
```


## Process the data

We add a column TYPE which tells us if the data comes from kinbank or if it is simulated, and we delete the first two columns because I am an idiot who ran my include_headings function too many times.

```{r}
gender_data <- gender_data %>%
  mutate(
    
    TYPE = ifelse (grepl('frankenlanguage', LANGUAGE),
    'simulation',
    'kinbank'
  ))


gender_data$GEN_0 <- as.numeric(gender_data$GEN_0)
gender_data$GEN_1 <- as.numeric(gender_data$GEN_1)


gender_data
```

And then we save gender_data to a new CSV file.

```{r}
write.csv(gender_data, '../../data/processed/all_gender_proportions.csv')
```

Now, let's split `gender_data` into two dataframes - `kinbank` and `simulations`.

```{r}
simulations <- filter(gender_data,TYPE == 'simulation')
kinbank <- filter(gender_data,TYPE == 'kinbank')
```

And save those too.

```{r}
write.csv(kinbank, '../../data/processed/kinbank_gender_proportions.csv')
write.csv(simulations, '../../data/processed/simulated_gender_proportions.csv')

```

# Analysis

We need the means of each SET of data - so all 1000 simulations, plus the kinbank data.

```{r}
simulation_means <- group_by(simulations,SET) %>%
  summarise(
    MEAN0 = mean(GEN_0),
    MEAN1 = mean(GEN_1)
  )

kinbank_mean <- group_by(kinbank,SET) %>%
  summarise(
    MEAN0 = mean(GEN_0),
    MEAN1 = mean(GEN_1)
  )

simulation_means
kinbank_mean
```

To plot the data, let's first take a sample of the data.

```{r}
sim_prop_sample <- sample_n(simulations,1000, replace=FALSE)
kinbank_prop_sample <- sample_n(kinbank,1000,replace=FALSE)
```

And we'll plot some histograms, so we can sanity check that we've done the simulations right.

First, does the count of each proportion look the same across the sims and the kinbank data?

```{r}
rbind(sim_prop_sample,kinbank_prop_sample) %>%
  ggplot() +
  geom_histogram(aes(GEN_0,fill=TYPE),position='dodge')

```

What about the mean proportions? This one is slightly concerning:

```{r}
simulation_means %>%
  ggplot(aes(MEAN0)) +
  geom_histogram(fill = 'lightblue', color = 'navy', alpha = 1) +
  geom_vline(aes(xintercept = 0.5868), color="navy", linetype="dashed", size=0.5) +
  labs(
    title = 'Mean proportion of gender distinctions in Ego generation',
    color = legend) +
  xlab('GEN_0 means')
```
Same plots for Ego's parents' generation:


```{r}
rbind(sim_prop_sample,kinbank_prop_sample) %>%
  ggplot() +
  geom_histogram(aes(GEN_1,fill=TYPE),position='dodge')

```

Oddly satisfying:

```{r}
simulation_means %>%
  ggplot(aes(MEAN1)) +
  geom_histogram(fill = 'lightblue', color = 'navy', alpha = 1) +
  geom_vline(aes(xintercept = 0.9595), color="navy", linetype="dashed", size=0.5) +
  labs(
    title = 'Mean proportion of gender distinctions in Ego parent generation',
    color = legend) +
  xlab('GEN_1 means')
```

Looks like the simulations converge to the real data! Phew!

## Real serious plots

Plot the data by generation:

```{r}
rbind(sim_prop_sample,kinbank_prop_sample) %>%
  ggplot(aes(x=GEN_0, y=GEN_1,color=TYPE)) +
  geom_jitter(alpha=0.2,width = 0.2) +
  xlim(0,1) +
  ylim(0,1) +
  geom_smooth(method=lm)
```

## Super serious analysis

### Spearman's doodad.

```{r}
cor.test(simulations$GEN_0,simulations$GEN_1, method = 'spearman')
```

```{r}
cor.test(kinbank$GEN_0,kinbank$GEN_1, method = 'spearman')

```

This is good, right? The correlation is way larger for kinbank, and the p-value is tiny, meaning that we wouldn't expect to get the
kinbank correlation by chance, but the simulation data is super likely to be chance (as we would expect because it is literally random).


