---
title: "symmetry scores"
author: "Maisy Hallam"
date: "2023-04-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # for data preprocessing
library(ggplot2) # for plotting
```

# Preprocessing the data

## Importing data

First, we import the three kinbank datasets.

```{r}
ge_kin <- read_csv('../../data/raw/gender/scores/kinbank_gender.csv')
ag_kin <- read_csv('../../data/raw/age/scores/kinbank_age.csv')
li_kin <- read_csv('../../data/raw/lineage/scores/kinbank_lineage.csv')
```

Next, we import the simulation data:

```{r}
ge <- list.files(path="../../data/raw/gender/scores/",pattern="set*",full.names=TRUE)
ge_sims <- do.call(rbind, lapply(ge, function(x) read.csv(x)))

ag <- list.files(path="../../data/raw/age/scores/",pattern="set*",full.names=TRUE)
ag_sims <- do.call(rbind, lapply(ag, function(x) read.csv(x)))

li <- list.files(path="../../data/raw/lineage/scores/",pattern="set*",full.names=TRUE)
li_sims <- do.call(rbind, lapply(li, function(x) read.csv(x)))
```

## Preprocessing the data

Make sure that the `feature` column is a factor.

```{r}
ge_kin$GENDER <- as.factor(ge_kin$GENDER)
ge_sims$GENDER <- as.factor(ge_sims$GENDER)

ag_kin$AGE <- as.factor(ag_kin$AGE)
ag_sims$AGE <- as.factor(ag_sims$AGE)

li_kin$LINEAGE <- as.factor(li_kin$LINEAGE)
li_sims$LINEAGE <- as.factor(li_sims$LINEAGE)
```


# Analysis

Let's identify the mean number of `[1,1]` scores for each feature in the simulation data.

```{r}
ge_sims_11<- group_by(ge_sims,SET) %>%
  summarise(
    mean_symmetrical = mean(GENDER == '[1, 1]')
  )

ag_sims_11 <- group_by(ag_sims,SET) %>%
  summarise(
    mean_symmetrical = mean(AGE == '[1, 1]')
  )

li_sims_11 <- group_by(li_sims,SET) %>%
  summarise(
    mean_symmetrical = mean(LINEAGE == '[1, 1]')
  )
```

And in the kinbank data.

```{r}
ge_kin_11 <- ge_kin %>%
  summarise(
    mean_symmetrical = mean(GENDER == '[1, 1]')
  )

ag_kin_11 <- ag_kin %>%
  summarise(
    mean_symmetrical = mean(AGE == '[1, 1]')
  )

li_kin_11 <- li_kin %>%
  summarise(
    mean_symmetrical = mean(LINEAGE == '[1, 1]')
  ) 

ge_kin_11
ag_kin_11
li_kin_11
```

## Plots

Let's plot the distribution of average 1,1 scores in the simulations, with the kinbank mean as a dashed line.

```{r}
ge_11 %>%
  ggplot(aes(mean_symmetrical)) +
  geom_density(fill = 'lightblue', alpha = 1) +
  geom_vline(aes(xintercept = 0.7684), color="navy", linetype="dashed", size=0.5) +
  labs(
    title = 'Mean proportion of 1,1 gender symmetry scores',
    color = legend) +
  xlab('Proportion of 1,1 gender scores') +
  xlim(0.6,0.9)
```

```{r}
ag_11 %>%
  ggplot(aes(mean_symmetrical)) +
  geom_density(fill = 'orange', alpha = 1) +
  geom_vline(aes(xintercept = 0.1373), color="brown", linetype="dashed", size=0.5) +
  labs(
    title = 'Mean proportion of 1,1 age symmetry scores',
    color = legend) +
  xlab('Proportion of 1,1 age scores') +
  xlim(0,0.25)
```

```{r}
li_11 %>%
  ggplot(aes(mean_symmetrical)) +
  geom_density(fill = 'lightblue', alpha = 1) +
  geom_vline(aes(xintercept = 0.4250), color="navy", linetype="dashed", size=0.5) +
  labs(
    title = 'Mean proportion of 1,1 lineage symmetry scores',
    color = legend) +
  xlab('Proportion of 1,1 lineage scores') +
  xlim(0.2,0.5)
```


