---
title: "austronesian-ics"
author: "Maisy Hallam"
date: "2023-07-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

##Import the data

```{r}
# sims <- read_csv('../../../data/raw/ics_sims.csv')
# kb <- read_csv('../../../data/raw/ics_real.csv')
```

Combine these dataframes. First, repeat all values of kb 1000 times.

```{r}
# kb_repeat <- kb %>% slice(rep(seq_len(n()), 1000))
# 
# kb_repeat <- arrange(kb_repeat, code)
# sims <- arrange(sims,code)
```

Then add the kb mutual_information column to sims as `real_mi`.

```{r}
# all <- sims %>% mutate(
#   real_mi = kb_repeat$mutual_information
# )
# 
# table(all$language_family)
```

## Sample the data

Take a random sample of languages.

```{r}
# languages <- sample(kb_filtered$code,12)
# 
# languages
```

And subset `all` by this sample.

```{r}
# sample <- subset(all, code %in% languages)
# 
# table(sample$code)
```

## Plot real values within simulated distributions

```{r}
# sample %>%
#   ggplot(aes(mutual_information,fill=language)) +
#   geom_density(alpha = 0.5) +
#   geom_vline(aes(xintercept = real_mi),linetype = 'dashed') +
#   facet_wrap(~language)
```

Let's do this again with our filtered dataset, to see if there's a difference.

```{r}
sims <- read_csv('../../../data/raw/ics_sims_filtered.csv')
kb <- read_csv('../../../data/raw/ics_real_filtered.csv')
```

Combine these dataframes. First, repeat all values of kb 1000 times.

```{r}
kb_repeat <- kb %>% slice(rep(seq_len(n()), 1000))

kb_repeat <- arrange(kb_repeat, code)
sims <- arrange(sims,code)
```

Then add the kb mutual_information column to sims as `real_mi`.

```{r}
all <- sims %>% mutate(
  real_mi = kb_repeat$mutual_information
)

all
```
Create a sample of languages:

```{r}
languages <- sample(all$code,12)

languages


```

Using the same sample of languages, plot this data:

```{r}
sample <- subset(all, code %in% languages)

sample
```

```{r}
sample %>%
  ggplot(aes(x=mutual_information,fill=language)) +
  geom_density(alpha = 0.5) +
  geom_vline(aes(xintercept = real_mi),linetype = 'dashed') +
  facet_wrap(~language)
```
Any difference between language families?

Take a random sample of languages (from kb_filtered to ensure they exist in both datasets)

```{r}
ie_languages <- sample(filter(all,language_family == 'Indo-European')$code,12)
au_languages <- sample(filter(all,language_family == 'Austronesian')$code,12)

ie_languages
au_languages
```

And subset `all` by this sample.

```{r}
ie_sample <- subset(all, code %in% ie_languages)
au_sample <- subset(all, code %in% au_languages)


```


```{r}
ie_sample %>%
  ggplot(aes(mutual_information,fill=language)) +
  geom_density(alpha = 0.5) +
  geom_vline(aes(xintercept = real_mi),linetype = 'dashed') +
  facet_wrap(~language)
```

```{r}
au_sample %>%
  ggplot(aes(mutual_information,fill=language)) +
  geom_density(alpha = 0.5) +
  geom_vline(aes(xintercept = real_mi),linetype = 'dashed') +
  facet_wrap(~language)
```
