---
title: "ics-z-score"
author: "Maisy Hallam"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
```

```{r}
theme_set(theme_bw())

# theme_update(text = element_text(family = "Fira Sans", size=20))

# axis.title.y = element_text(angle=90, vjust=0.5, hjust = 0))
```

## Import the data

```{r}
sims <- read_csv('../../data/raw/simulated-kinship-mi-data.csv')
kb <- read_csv('../../data/raw/full-kinship-mi-data.csv')
```

What does the data look like for English?

```{r}
sims %>% filter(code == 'stan1293')

kb %>% filter(code == 'stan1293')
```


Get the mean mutual information score across simulations for each language.

```{r}
mi <- sims %>% 
  group_by(code) %>% 
  summarise(
    mean = mean(mutual_information), 
    sd = sd(mutual_information))
```

Order kb in alphabetical order.

```{r}
# mean_mi$real_value = kb$mutual_information

filter(mi, code == 'stan1293')


kb <- arrange(kb, code)
mi <- arrange(mi,code)

mi <- mi %>% mutate(
  real_value = kb$mutual_information,
  language = kb$language
)

filter(mi,language == 'Tagalog')
```


Add the z-score for each language - the real MI value minus the mean MI value of the simulations, divided by the SD of the simulations.

```{r}
mi$mi_z <- ((mi$real_value - mi$mean) / mi$sd)

mi
```

Repeat kb 1000 times.

```{r}
kb_repeat <- kb %>% slice(rep(seq_len(n()), 1000))

kb_repeat

kb_repeat = kb_repeat[order(kb_repeat$code),]
```

```{r}
kb_repeat
```

Combine kb and sims

```{r}

sims <- arrange(sims, code)
kb_repeat <- arrange(kb_repeat, code)
all <- sims
all$real_value <- kb_repeat$mutual_information

filter(all, code == 'stan1293')
```

## Sample languages

Take a random sample of languages.

```{r}
ie <- c(filter(kb, mutual_information > 0, language_family == 'Indo-European')$code %>% sample(3), 'stan1293')
au <- c(filter(kb, mutual_information > 0, language_family == 'Austronesian')$code %>% sample(3), 'taga1270')
pn <- c(filter(kb, mutual_information > 0, language_family == 'Pama-Nyungan')$code %>% sample(3), 'warl1254')

ie
```

```{r}
# original sample
ie <- c('Dari','Hindi','Russian','English')
au <- c("Tanampedagi_Taje","Tikopia","Watut", "Tagalog")
pn <- c("Bunganditj","Ngiyambaa","Wajarri", "Warlpiri")
```

And subset `mi_repeat` by this sample.

```{r}
ie_sample <- subset(all, code %in% ie)
au_sample <- subset(all, code %in% au)
pn_sample <- subset(all, code %in% pn)

table(au_sample$language)
```

```{r}
ie_sample <- subset(all, language %in% ie)
au_sample <- subset(all,language %in% au)
pn_sample <- subset(all,language %in% pn)

table(au_sample$language)

ie_sample <- filter(ie_sample,code != 'russ1263a')
au_sample <- filter(au_sample,code != 'tiko1237')
pn_sample <- filter(pn_sample,code != 'warl1254a')


```

## Plot real values within simulated distributions

```{r}
ie_sample$language = factor(ie_sample$language, levels=c("Dari","Hindi","Russian", "English"))

ie_plot <- ie_sample %>%
  ggplot(aes(mutual_information)) +
  geom_histogram(colour = '#d95f02', fill = '#d95f02') +
  geom_vline(aes(xintercept = real_value),linetype = 4) +
  ggtitle('Indo-European') +
  xlab('') +
  facet_grid(~language) +
  theme(
    panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    # panel.grid.major = element_blank(), #remove major gridlines
    # panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent'), #transparent legend bg
    legend.box.background = element_rect(fill='transparent') #transparent legend panel
  )

ie_plot

ggsave('../../plots/ie_plot.png', ie_plot, bg = 'transparent',width = 9, height = 3)

```

```{r}
au_sample$language = factor(au_sample$language, levels=c("Tanampedagi_Taje","Tikopia","Watut", "Tagalog"))

au_plot <- au_sample %>%
  ggplot(aes(mutual_information)) +
  geom_histogram(colour = '#7570b3', fill = '#7570b3') +
  geom_vline(aes(xintercept = real_value),linetype = 4) +
  ggtitle('Austronesian') +
  xlab('') +
  facet_grid(~language) +
  theme(
    panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    # panel.grid.major = element_blank(), #remove major gridlines
    # panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent'), #transparent legend bg
    legend.box.background = element_rect(fill='transparent') #transparent legend panel
  )

au_plot

ggsave('../../plots/au_plot.png', au_plot, bg = 'transparent',width = 9, height = 3)

```

```{r}
pn_sample$language = factor(pn_sample$language, levels=c("Bunganditj","Ngiyambaa","Wajarri", "Warlpiri"))

pn_plot <- pn_sample %>%
  ggplot(aes(mutual_information)) +
  geom_histogram(colour = '#1b9e77', fill = '#1b9e77') +
  geom_vline(aes(xintercept = real_value),linetype = 4) +
  ggtitle('Pama-Nyungan') +
  xlab('Mutual information') +
  facet_grid(~language) +
  theme(
    panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    # panel.grid.major = element_blank(), #remove major gridlines
    # panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent'), #transparent legend bg
    legend.box.background = element_rect(fill='transparent') #transparent legend panel
  )

pn_plot

ggsave('../../plots/pn_plot.png', pn_plot, bg = 'transparent',width = 9, height = 3)
```

```{r}
grid_plot <- ie_plot + au_plot + pn_plot +
  plot_layout(nrow = 3)
  # plot_annotation(title = 'Distribution of mutual information for simulated kinship systems')

grid_plot

ggsave('../../plots/sample_of_languages_tsp.png', width = 9, height = 9, bg='transparent')
```


```{r}
all_long <- rbind(sims,kb)

filter(all_long,simulation == 'N')
```

```{r}
filter(all_long,mutual_information > 0) %>%
  ggplot(aes(mutual_information, fill = simulation)) +
  scale_fill_manual(values = c('#d95f02','#7570b3'),
                    name="data type:",
                    labels=c("natural languages", "simulations")) +
  geom_density(alpha = 0.5) +
  # ggtitle('Distribution of mutual information for real and simulated kinship systems') +
  xlab('Mutual information (bits)') +
  theme_minimal() + 
  theme(
    axis.line.x = element_line(color = "black"),
    axis.line.y = element_line(color = 'black'),
    legend.position = c(0.7, 0.8))

ggsave('../../plots/all_sims_distributions_tsp.png',height=5,width=5,bg='transparent')
```

```{r}

hindi_plot <- filter(all,language == 'Hindi') %>%
  ggplot(aes(mutual_information)) +
  geom_histogram(colour = '#d95f02', fill = '#d95f02') +
  geom_vline(aes(xintercept = real_value),linetype = 4,size=1) +
  # ggtitle('Indo-European') +
  xlab('Mutual information (bits)')
  # facet_grid(~language) +
  # theme_bw() +
  # theme(
    # panel.background = element_rect(fill='transparent'), #transparent panel bg
    # plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    # # panel.grid.major = element_blank(), #remove major gridlines
    # # panel.grid.minor = element_blank(), #remove minor gridlines
    # legend.background = element_rect(fill='transparent'), #transparent legend bg
    # legend.box.background = element_rect(fill='transparent') #transparent legend panel

hindi_plot
ggsave('hindi_plot.png',hindi_plot, width=5,height=5)

```

```{r}
tagalog_plot <- filter(all,language == 'Tagalog') %>%
  ggplot(aes(mutual_information)) +
  geom_histogram(colour = '#d95f02', fill = '#d95f02') +
  geom_vline(aes(xintercept = real_value),linetype = 4,size=1)+
  # ggtitle('Indo-European') +
  xlab('Mutual information (bits)') +
  ylab('')
  # facet_grid(~language) +
  # theme_bw() +
  # theme(
    # panel.background = element_rect(fill='transparent'), #transparent panel bg
    # plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    # # panel.grid.major = element_blank(), #remove major gridlines
    # # panel.grid.minor = element_blank(), #remove minor gridlines
    # legend.background = element_rect(fill='transparent'), #transparent legend bg
    # legend.box.background = element_rect(fill='transparent') #transparent legend panel

ggsave('tagalog_plot.png',tagalog_plot, width=5,height=5)
```

```{r}
hind_tag_plot <- hindi_plot + tagalog_plot +
  plot_layout(widths = c(2,2))
  # plot_annotation(title = 'Distribution of mutual information for simulated kinship systems')

hind_tag_plot

ggsave('../../../sample_of_languages_2.png', width = 9, height = 9, bg='transparent')

```

```{r}
hind_sund_plot <- filter(all, language == 'English' | language == 'Hindi' | language == 'Sundanese' | code == 'warl1254') %>%
  ggplot(aes(mutual_information)) +
  geom_histogram(colour = '#d95f02', fill = '#d95f02') +
  geom_vline(aes(xintercept = real_value),linetype = 2,linewidth=1) +
  xlab('Mutual information (bits)') +
  facet_wrap(~language,ncol =2)

hind_sund_plot

ggsave('hind_sund_plot.png',hind_sund_plot,height=5,width=5)
```

```{r}
table(filter(all,language_family == 'Pama-Nyungan')$code)
```

